The Emergent Simplicity of Galaxy Size

### 2017-09-15 paper/galsize_paper.tex
Stripped away remaining obsolete text.

%In \S\ref{subsec:tests1} we show comparisons between the galaxy size model described in \S\ref{sec:model} and our SDSS sample. We identify the key ingredients that determine the characteristic $\rhalf-$dependence of galaxy clustering in \S\ref{subsec:tests2}. In so doing, we demonstrate the sensitivity of galaxy clustering measurements to the underlying model assumptions, establishing the success of our model as non-trivial.

@@

We measure $\wproj(\rproj)$ separately for large and small subsamples for four different $\mstar$ thresholds, $\mstar>10^{9.75}\msun,$ $\mstar>10^{10.25}\msun,$ $\mstar>10^{10.75}\msun,$ and $\mstar>10^{11.25}\msun.$ We make the same measurements for each volume-limited $\mstar-$threshold sample {\em without} splitting on size, giving us measurements $\wpall, \wplarge,$ and $\wpsmall$ for each threshold sample. This allow us to compute the ratio $(\wplarge-\wpsmall)/\wpall,$ which we refer to as {\em the $\rhalf$ clustering ratio}. These ratios are the measurements appearing on the y-axis in each panel of Figure \ref{fig:clustering_ratio_upshot}. Points with error bars show SDSS measurements, solid curves show the clustering ratios of model galaxies. Before unpacking the information contained in these clustering measurements, we stress that the good agreement shown between model and data in Figure~\ref{fig:clustering_ratio_upshot} is a genuine model prediction, since the model parameters were taken directly from \citet{kravtsov13}, which were fit to the one-point measurements in Fig.~\ref{fig:scatter_plot}, whereas the two-point measurements appearing in Figure \ref{fig:clustering_ratio_upshot} have heretofore never been measured.

@@

The salient feature of the clustering ratio measurements is that they are negative: small galaxies cluster more strongly than large galaxies of the same stellar mass, a new result. This feature also holds true for model galaxies. This result may be surprising, since $\rhalf\propto\rvir,$ halo mass $\rvir\propto\mhalo^{1/3},$ and clustering strength increases with $\mvir.$ Based on this simple argument, one would expect the opposite trend to the measurements shown here.

The resolution to this puzzle is illustrated in Figure~\ref{fig:bt_censat}, which shows the PDF of $\bt$ for halos of the same mass $\mhalo,$ defined as $\mpeak,$ the peak mass ever attained through the history of the (sub)halo, so that host halos and subhalos can be treated on equal footing. In the model, central galaxies are diskier than satellites, a feature that is inherited from the method we use to map $\bt$ to model galaxies.

In light of this explanation, it is natural to ask whether this simple feature {\em alone} is all that is needed for any model of $\rhalf$ to achieve this level of success at predicting galaxy clustering on small and large scales. We address this question in \S~\ref{subsec:tests2} below, finding that the answer is no: the reasonably correct magnitude and scale-dependence of the clustering ratios predicted by our model is a non-trivial result that places tight constraints on the post-infall physics of satellite galaxies.

\subsection{Sensitivity of Clustering to Satellite Profile Evolution}
\label{subsec:tests2}

In this section we explore how post-infall changes in satellite $\rhalf$ manifest in $\wproj(\rproj).$ To do so, we construct an extension our fiducial model with a simple additional ingredient for post-infall size evolution of satellites.

In the fiducial model described in \S~\ref{sec:model}, recall that in the power law a scaling relations (Eq.~\ref{eq:fiducial_model}) the halo radius $\rvir$ used for satellite galaxies is taken to be the value at the time of infall. The simple interpretation of this assumption is that, in a statistical sense, the size of satellite galaxies is determined by its history as a central galaxy, and that post-infall physics leaves no distinct imprint on satellite $\rhalf.$

Here we suppose that this assumption is violated, and that $\rhalf$ decreases after infall according to $(\mvir/\macc)^{1/3}.$ That is, in this alternative formulation, rather than using $\rvir$ at the time of infall for satellites in the power law scaling relations, we instead use $\rvir(\mvir/\macc)^{1/3}.$ This simple toy model attempts supposes that the same physical processes leading to halo mass loss are also responsible for a post-infall decrease in satellite size.

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/scrambled_size_clustering_ratios.pdf}
\caption{
{\bf Origin of the $\rhalf-$dependence of clustering.}
Here we compare our fiducial model, in which satellite galaxy size is set by $\rvir$ at the time of infall, to an alternative model analogous to \citet{watson_etal12} in which satellite sizes contract in proportion to $(\mvir/\macc)^{1/3}.$ The large differences between solid and dashed curves in the top panels show that the $\rhalf-$dependence of galaxy clustering ratios is highly sensitive to the post-infall evolution of satellite galaxy profiles. The successful prediction of our fiducial model, in which satellite galaxies neither contract nor puff up after infall, places tight constraints on satellite-specific physical processes, which must be either negligible or conspiratorially produce little-to-no size change after accretion.
}
\label{fig:satellites}
\end{figure}
%-----------------------------------------------------------------------------------------------------

Figure~\ref{fig:satellites} compares the clustering ratios of our fiducial model (solid curves) to the clustering ratios of this alternative model (dashed curves). The difference between the small-scale clustering of the two models is stark: the assumption that $\rhalf$ decreases after infall leaves a strong imprint on the $\rhalf-$dependence of $\wproj(\rproj),$ such that small galaxies cluster {\em much} more strongly relative to large galaxies, particularly for disk-dominated systems with $\mstar\approx10^{10}\msun.$

The large differences between the solid and dashed curves in Fig.~\ref{fig:satellites} establishes that the largely successful prediction for the clustering ratios fiducial model is a non-trivial: $\wproj(\rproj)$ is indeed providing good constraining power on the  assumptions underlying our profile modeling and not simply our morphology modeling, c.f., \S\ref{subsubsec:random_bt_model} and \S\ref{sec:model}. We refer the reader to \S\ref{subsec:satellite_discussion} for further discussion of the physical implications of this result.

>>>>

In \S\ref{subsec:predictions} we show comparisons between the galaxy size model described in \S\ref{sec:model} and our SDSS sample. We identify the key ingredients that determine the characteristic $\rhalf-$dependence of galaxy clustering in \S\ref{subsec:subsec:shuffle_tests}. In so doing, we demonstrate the sensitivity of galaxy clustering measurements to the underlying model assumptions, establishing the success of our model as non-trivial.

\subsection{Testing Model Predictions}
\label{subsec:predictions}

@@

We measure $\wproj(\rproj)$ separately for large and small subsamples for four different $\mstar$ thresholds, $\mstar>10^{9.75}\msun,$ $\mstar>10^{10.25}\msun,$ $\mstar>10^{10.75}\msun,$ and $\mstar>10^{11.25}\msun.$ We make the same measurements for each volume-limited $\mstar-$threshold sample {\em without} splitting on size, giving us measurements $\wpall, \wplarge,$ and $\wpsmall$ for each threshold sample. This allow us to compute the ratio $(\wplarge-\wpsmall)/\wpall,$ which we refer to as {\em the $\rhalf$ clustering ratio}. These ratios are the measurements appearing on the y-axis in each panel of Figure \ref{fig:clustering_ratio_upshot}. Points with error bars show SDSS measurements, solid curves show the clustering ratios of model galaxies as predicted by the model described in \S\ref{sec:model}. Before unpacking the information contained in these clustering measurements, we stress that the good agreement shown between model and data in Figure~\ref{fig:clustering_ratio_upshot} is a genuine model prediction, since the model parameters were taken directly from \citet{kravtsov13}, which were fit to the one-point measurements in Fig.~\ref{fig:scatter_plot}, whereas the two-point measurements appearing in Figure \ref{fig:clustering_ratio_upshot} have heretofore never been measured.

@@

The salient feature of the clustering ratio measurements is that they are negative: small galaxies cluster more strongly than large galaxies of the same stellar mass, a new result. This feature also holds true for model galaxies. This result may be surprising, since $\rhalf\propto\rvir,$ halo mass $\rvir\propto\mhalo^{1/3},$ and clustering strength increases with $\mvir.$ Based on this simple argument, one would expect the opposite trend to the measurements shown here. We present a resolution to this puzzle in the following section.

\subsection{Origin of the size-dependence of galaxy clustering}
\label{subsec:shuffle_tests}

We can understand the origin of the shape of the clustering strength ratios shown in Figure~\ref{fig:clustering_ratio_upshot} through a series of straightforward ``shuffle tests".

% shown in Figure \ref{fig:shuffle}.

%In light of this explanation, it is natural to ask whether this simple feature {\em alone} is all that is needed for any model of $\rhalf$ to achieve this level of success at predicting galaxy clustering on small and large scales. We address this question in \S~\ref{subsec:tests2} below, finding that the answer is no: the reasonably correct magnitude and scale-dependence of the clustering ratios predicted by our model is a non-trivial result that places tight constraints on the post-infall physics of satellite galaxies.

We conclude this section by estimating how satellite mass stripping manifests in $\wproj(\rproj).$ To do so, we construct an extension our fiducial model with a simple additional ingredient for post-infall evolution of satellites mass and size.

%In the fiducial model described in \S~\ref{sec:model}, recall that in the power law a scaling relations (Eq.~\ref{eq:fiducial_model}) the halo radius $\rvir$ used for satellite galaxies is taken to be the value at the time of infall. The simple interpretation of this assumption is that, in a statistical sense, the size of satellite galaxies is determined by its history as a central galaxy, and that post-infall physics leaves no distinct imprint on satellite $\rhalf.$

%Here we suppose that this assumption is violated, and that $\rhalf$ decreases after infall according to $(\mvir/\macc)^{1/3}.$ That is, in this alternative formulation, rather than using $\rvir$ at the time of infall for satellites in the power law scaling relations, we instead use $\rvir(\mvir/\macc)^{1/3}.$ This simple toy model attempts supposes that the same physical processes leading to halo mass loss are also responsible for a post-infall decrease in satellite size.

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/scrambled_size_clustering_ratios.pdf}
\caption{
{\bf Origin of the $\rhalf-$dependence of clustering.}
Here we compare our fiducial model, in which satellite galaxy size is set by $\rvir$ at the time of infall, to a set of alternative models created by shuffling the sizes of various subsamples of the fiducial mock. 
}
\label{fig:shuffle}
\end{figure}
%-----------------------------------------------------------------------------------------------------

%Figure~\ref{fig:satellites} compares the clustering ratios of our fiducial model (solid curves) to the clustering ratios of this alternative model (dashed curves). The difference between the small-scale clustering of the two models is stark: the assumption that $\rhalf$ decreases after infall leaves a strong imprint on the $\rhalf-$dependence of $\wproj(\rproj),$ such that small galaxies cluster {\em much} more strongly relative to large galaxies, particularly for disk-dominated systems with $\mstar\approx10^{10}\msun.$

%The large differences between the solid and dashed curves in Fig.~\ref{fig:satellites} establishes that the largely successful prediction for the clustering ratios fiducial model is a non-trivial: $\wproj(\rproj)$ is indeed providing good constraining power on the  assumptions underlying our profile modeling and not simply our morphology modeling, c.f., \S\ref{subsubsec:random_bt_model} and \S\ref{sec:model}. We refer the reader to \S\ref{subsec:satellite_discussion} for further discussion of the physical implications of this result.
